Binarity's Biases on Transit Survey Occurrence Rates

### 2018-02-23 paper/main.tex
add git commit
jnw intro edits

\shortauthors{Bouma, Masuda, Winn}
\shorttitle{Binarity's Biases}

@@

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%

Wide-field surveys for transiting planets, such as the NASA {\it
Kepler} and {\it TESS} missions, are usually conducted without knowing
which stars are binaries. Unresolved and unrecognized binary stars
give rise to systematic errors in planet occurrence rates, including
misclassified planets and miscounts in the number of searched stars.
The individual errors can have different signs, making it difficult to
anticipate the net effect on inferred occurrence rates. Here we use
simplified models of signal-to-noise limited transit surveys to try
and clarify the situation. We derive a formula for the apparent
occurrence rate density measured by an observer who falsely assumes
all stars are single. The formula depends on the binary fraction; the
mass function of the secondary stars; and the true occurrence of
planets around primaries, secondaries, and single stars. It also takes
into account the Malmquist bias by which binaries are over-represented
in flux-limited samples. Application of the formula to an idealized
{\it Kepler}-like survey shows that for planets larger than
$2r_\oplus$, the net systematic error is of order 10\%.  One
consequence is that unrecognized binaries cannot account for the
apparent discrepancy between hot Jupiter occurrence rates measured by
transit and RV surveys.  For smaller planets the errors are
potentially larger: the occurrence of Earth-sized planets could be
overestimated by as much as a factor of two. We also show that if
high-resolution imaging reveals a transit host star to be a binary,
the planet is usually more likely to orbit the primary star than the
secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

One of the goals of exoplanetary science is to establish how common,
or rare, are planets of various types.  Knowledge of planet occurrence
rates is helpful for inspiring and testing theories of planet
formation, designing the next generation of planet-finding surveys,
and simply satisfying our curiosity about other worlds.  One method
for measuring occurrence rates is to monitor the brightnesses of many
stars over a wide field, seeking evidence for planetary transits.
Measuring occurrence rates using this method was the highest priority
of the NASA {\it Kepler} mission.  Great strides have been made in the
analysis of {\it Kepler} data, including progress towards measuring
the fraction of Sun-like stars that harbor Earth-like planets~\citep{
  youdin_exoplanet_2011,petigura_prevalence_2013,dong_fast_2013,
  foreman-mackey_exoplanet_2014,burke_terrestrial_2015}

A lingering concern in occurrence rate studies is that in most cases,
investigators have assumed that all of the sources of light that were
monitored are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the point sources seen in a transit survey are
unresolved multiple-star systems, especially binaries.  Unrecognized
binaries cause numerous systematic errors in the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, in a binary there are two opportunities to detect
transiting planets, which could increase the overall number of
detections.

At the outset of this study it was not clear to us whether the neglect
of binaries was a serious issue at all, or even whether the net effect
of all the errors is positive or negative.  The goals of this study
were to provide a framework for dealing with these issues, and to
calculate a leading-order estimate for the net systematic effects.  In
this spirit, our models are idealized and analytic, and we do not
attempt a detailed correction of the results from {\it Kepler} or any
other real transit survey.

This paper is organized as follows.  The next section enumerates the
various errors that arise from unrecognized binaries.  Then in
Section~\ref{sec:simplest}, we develop an idealized model of a transit
survey in which all planets have identical properties, and all stars
are identical except that some fraction are in binary systems.  This
simple model motivates the derivation of a general formula, given in
Section~\ref{sec:general_formula}, that allows for more realistic
stellar and planetary populations.  We use this formula in
Section~\ref{sec:more_complicated} to explore more complicated but
nevertheless still analytic models.  In Section~\ref{sec:discussion}
we discuss the errors due to unrecognized binaries for specific cases
of current interest: the occurrence of Earth-like planets; the
apparent discrepancy between hot Jupiter occurrence rates in different
surveys; and the shape of the ``evaporation valley'' in the planet
radius distribution that was brought to light by
\citet{fulton_california-_2017}.  We review all the results in
Section~\ref{sec:conclusion}.

@@

Though the details of the planet radius distribution $\Gamma_i(r)$ at
$r<2r_\oplus$ are currently an active topic of research, by making
plausible assumptions we can investigate how neglecting binarity might
bias measurements in this regime.  For instance, consider a broken
power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2r_\oplus \\
        {\rm constant} & \text{for } r\leq2r_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
At apparent radii $r\a > 2r_\oplus$, the apparent rate density in
this model is the same as that in
Section~\ref{sub:powerlaw_varying_binaries}.  At smaller apparaent
radii the equations are tedious, but still tractable.
For simplicity, we insert Equation~\ref{eq:model3_radius_distribution}
into Equation~\ref{eq:general_Gamma_a}, and integrate using a computer
program. We refer the interested reader to our online
implementation\footnote{\url{https://github.com/lgbouma/binary_biases}}.
The output, shown in Figure~\ref{fig:occ_rate_model_3_log}, is
validated using analytic predictions in the $r\a > 2r_\oplus$ and the
$r\a < 2r_\oplus/\sqrt{2}$ regimes.  As before, $L\propto M^\alpha
\propto R^\alpha$, and $\alpha=3.5$, $\beta=0$, $\gamma=-2.92$.

>>>>

\shortauthors{Bouma, Masuda, Winn}
\shorttitle{Binarity and planet occurrence rates}

@@

\title{ Binarity's Biases on Transit Survey Occurrence Rates }
%
\correspondingauthor{L. Bouma}
\email{luke@astro.princeton.edu}
%
\author[0000-0002-0514-5538]{L. G. Bouma}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0003-1298-9699]{K. Masuda}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
\author[0000-0002-4265-047X]{J. N. Winn}
\affiliation{
    Department of Astrophysical Sciences,
    Princeton University,
    4 Ivy Lane, Princeton, NJ 08540, USA}
%
%
\begin{abstract}
%
Wide-field surveys for transiting planets, such as the NASA {\it
  Kepler} and {\it TESS} missions, are usually conducted without
knowing which stars have binary companions. Unresolved and
unrecognized binaries give rise to systematic errors in planet
occurrence rates, including misclassified planets and mistakes in
completeness corrections.  The individual errors can have different
signs, making it difficult to anticipate the net effect on inferred
occurrence rates. Here we use simplified models of signal-to-noise
limited transit surveys to try and clarify the situation. We derive a
formula for the apparent occurrence rate density measured by an
observer who falsely assumes all stars are single. The formula depends
on the binary fraction; the mass function of the secondary stars; and
the true occurrence of planets around primaries, secondaries, and
single stars. It also takes into account the Malmquist bias by which
binaries are over-represented in flux-limited samples. Application of
the formula to an idealized {\it Kepler}-like survey shows that for
planets larger than 2\,$R_\oplus$, the net systematic error is of
order 5\%.  In particular, unrecognized binaries are unlikely to be
the reason for the apparent discrepancies between hot Jupiter
occurrence rates measured in different surveys.  For smaller planets
the size the errors are potentially larger: the occurrence of
Earth-sized planets could be overestimated by as much as 50\%. We also
show that whenever high-resolution imaging reveals a transit host star
to be a binary, the planet is usually more likely to orbit the primary
star than the secondary star.
%
\end{abstract}
%
\keywords{
    methods: data analysis ---
    planets and satellites: detection ---
    surveys}
%
%

@@

One of the goals of exoplanetary science is to establish how common,
or rare, are planets of various types.  Knowledge of planet occurrence
rates is helpful for inspiring and testing theories of planet
formation, designing the next generation of planet-finding surveys,
and simply satisfying our curiosity.  One method for measuring
occurrence rates is to monitor the brightnesses of many stars over a
wide field, seeking evidence for planetary transits.  This was the
highest priority of the NASA {\it Kepler} mission.  Great strides have
been made in the analysis of {\it Kepler} data, including progress
towards measuring the fraction of Sun-like stars that harbor
Earth-like planets~\citep{
  youdin_exoplanet_2011,petigura_prevalence_2013,dong_fast_2013,
  foreman-mackey_exoplanet_2014,burke_terrestrial_2015}.

A lingering concern about these studies is that in most cases,
investigators have assumed that all of the targets in the survey
are single stars~\citep[\textit{e.g.},][]{
  howard_planet_2012,fressin_false_2013,
  dressing_occurrence_2015,burke_terrestrial_2015}. 
In reality, many of the sources that are monitored in a transit survey are
unresolved multiple-star systems, mainly binaries.  Unrecognized
binaries cause numerous systematic errors in the planetary occurrence
rates.  For example, when there is a transiting planet around a star
in a binary, the additive constant light from the second star reduces
the fractional loss of light due to the planet.  This makes transit
signals harder to detect and lowers the number of detections.  On the
other hand, a binary system presents two opportunities to detect
transiting planets, which could increase the overall number of
detections.

At the outset of this study it was not clear to us whether the neglect
of binaries is a serious problem, or even whether the net effect of
the errors is positive or negative.  The goals of this study were to
provide a framework for dealing with these issues, and to gauge at
least the order of magnitude of the systematic errors.  In this
spirit, our models are idealized.  We do not attempt a detailed
correction of the results from {\it Kepler} or any other real transit
survey.  We took inspiration from the insightful analytic model for
transit surveys by~\citet{pepper_using_2003}.

This paper is organized as follows.  The next section enumerates the
various errors that arise from unrecognized binaries.  Then in
Section~\ref{sec:simplest}, we develop an idealized model of a transit
survey in which all planets have identical properties, and all stars
are identical except that some fraction are in binary systems.  This
simple model motivates the derivation of a general formula, given in
Section~\ref{sec:general_formula}, that allows for more realistic
stellar and planetary populations.  We use this formula in
Section~\ref{sec:more_complicated} to explore more realistic models.
We discuss the errors due to unrecognized binaries for specific cases
of current interest: the occurrence of Earth-like planets; the
apparent discrepancy between hot Jupiter occurrence rates measured in
different surveys; and the shape of the ``evaporation valley'' in the
planet radius distribution that was brought to light by
\citet{fulton_california-_2017}.  We summarize and discuss all the
results in Section~\ref{sec:discussion}.

@@

Though the details of the planet radius distribution $\Gamma_i(r)$ at
$r<2r_\oplus$ are currently an active topic of research, by making
plausible assumptions we can investigate how neglecting binarity might
bias measurements in this regime.  For instance, consider a broken
power-law:
\begin{align}
    f(r)
    &\propto
    \left.
    \begin{cases}
        r^\gamma & \text{for } r\geq 2r_\oplus \\
        {\rm constant} & \text{for } r\leq2r_\oplus.
    \end{cases}
    \right.
    \label{eq:model3_radius_distribution}
\end{align}
At apparent radii $r\a > 2r_\oplus$, the apparent rate density in
this model is the same as that in
Section~\ref{sub:powerlaw_varying_binaries}.  At smaller apparaent
radii the equations are tedious, but still tractable.
For simplicity, we insert Equation~\ref{eq:model3_radius_distribution}
into Equation~\ref{eq:general_Gamma_a}, and integrate using a computer
program. We refer the interested reader to our online
implementation\footnote{\url{https://github.com/lgbouma/binary_biases},
commit \texttt{6cb920c}}.
The output, shown in Figure~\ref{fig:occ_rate_model_3_log}, is
validated using analytic predictions in the $r\a > 2r_\oplus$ and the
$r\a < 2r_\oplus/\sqrt{2}$ regimes.  As before, $L\propto M^\alpha
\propto R^\alpha$, and $\alpha=3.5$, $\beta=0$, $\gamma=-2.92$.
